SubsidyExplorer - Step 2: Create Vessel List

### 2020-03-15 scripts/02_vessel_list.Rmd
good start on figure paper redo

title: "SubsidyExplorer - Step 1: Create Vessel List"
author: "Kat Millage, Vienna Saccomanno, Laura Lea Rubino, Christopher Costello"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
bibliography: "../pew-subsidies.bib"

@@

@@

# Path to cleaned subsidies data file
subsidy_dat_path <- here::here("data", "edited", "sumaila-subsidies-2018-tidy.csv")

# Path to cleaned landings data file
catch_dat_path <- here::here("data", "edited", "fisheries-area-2000-2017-tidy.csv")

# Bigquery project
bq_project <-  "ucsb-gfw"

# Do you want to make summary figures?
make_figures <- F
```

@@

# Directories for the output files generated by this script
results_dir <- here::here("results", "1-vessel-list/")
if (dir.exists(results_dir) == F) {
  dir.create(results_dir, recursive = T)
}

@@

@@

@@

  bq_table_delete(effort_table)
  effort_good_fishing_vessels_2018 <- 
    bq_project_query(bq_project, 
                    sql2, 
                    destination_table = effort_table, 
                    use_legacy_sql = FALSE, 
                    allowLargeResults = TRUE) %>%
     bq_table_download()

}else {
  bq_table_create(effort_table)
  effort_good_fishing_vessels_2018 <- 
    bq_project_query("world-fishing-827", 
                   query = sql2, 
                   destination_table = effort_table,
                   use_legacy_sql = FALSE, 
                   allowLargeResults = TRUE) %>%
    bq_table_download()
}

# Save
write_csv(effort_good_fishing_vessels_2018, paste0(results_dir, "effort_good_fishing_vessels_2018_raw.csv"))

# If BigQuery connection isn't working, or you don't have access, impuort from local file
effort_good_fishing_vessels_2018 <- read_csv(paste0(results_dir, "effort_good_fishing_vessels_2018_raw.csv"), col_types = "nnnnnccnnnccnccccccn")
```

>>>>

title: "SubsidyExplorer - Step 2: Create Vessel List"
author: "Kat Millage, Vienna Saccomanno, Laura Lea Rubino, Christopher Costello"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
bibliography: "../pew-subsidies.bib"

@@

It also creates a 5x5 degree version for use in making the figures for the paper, but this version is kept on BigQuery.

@@

# Path to cleaned subsidies data file
subsidy_dat_sumaila_path <- here::here("data", "edited", "sumaila_et_al_2019_subsidies_tidy.csv")

# Path to cleaned landings data file
#catch_dat_path <- here::here("data", "edited", "fisheries-area-2000-2017-tidy.csv")

# Bigquery project
bq_project <-  "emlab-gcp"

@@

# Directories for the output files generated by this script
results_dir <- here::here("results", "01-vessel-list/")
if (dir.exists(results_dir) == F) {
  dir.create(results_dir, recursive = T)
}

@@

## By 5x5 degree latititude/longitude

We first extract the number of hours spent fishing by each of our good fishing vessels in each 5x5 degree latitude/longitude bin from the raw fishing effort table (`world-fishing-827.gfw_research.pipe_v20190502_fishing`).

To identify valid fishing effort here, we apply the following filters: 
- The fishing activity must have occured during a valid segment as identified by GFW.
- The fishing activity must have occured during a segment with more than 10 positions that was not concurrently overlapping with another, longer, segment for the same MMSI.  
- The vessel performing the fishing activity must appear on the list of good fishing vessels created above

We then extract a vessel table with the the following attributes: 
- year (int): Year. Matches our data year specified above.
- lat_bin (int): 5 degree latitude bin, coordinates refer to the center of the bin
- lon_bin (int): 5 degree longitude bin, coordinates refer to the center of the bin
- fao_region (int): Major FAO statistical region
- fishing_hours_lat_lon: Total annual fishing hours in the specified lat/lon bin by a single vessel
- ssvid (chr): Unique vessel identification number given by GFW (MMSI number for most vessels)
- flag (chr): Best flag state of the vessel.

  # This subquery identifies good fishing vessels from our previous query
  good_vessels AS (
  SELECT
    *
  FROM
    `emlab-gcp.SubsidyExplorer.all_good_fishing_vessels`),

  # This subquery identifies vessel information that we want to keep
  vessel_info AS (
  SELECT
    ssvid,
    year,
    best.best_flag as flag,
    best.best_engine_power_kw as engine_power_kw
  FROM
    `world-fishing-827.gfw_research.vi_ssvid_byyear_v20190430`
  WHERE
    year = 2018),

  # This subquery gets all fishing effort by ssvid, eez, and fao region in 2018 cooresponding to good segments and good vessels
  fishing_by_lat_lon AS (
    SELECT
      EXTRACT(year FROM date) AS year,
      ssvid,
      FLOOR(lat*0.5)/0.5 + 2.5 lat_bin_center,
      FLOOR(lon*0.5)/0.5 + 2.5 lon_bin_center, 
      if(ARRAY_LENGTH(regions.major_fao)=0, "0", regions.major_fao[ordinal(1)]) AS fao_region,
      sum(hours) AS fishing_hours_lat_lon
    FROM
      `world-fishing-827.gfw_research.pipe_v20190502_fishing`
    WHERE
      date > timestamp("2018-01-01")
      AND date < timestamp("2018-12-31")
      AND ARRAY_LENGTH(regions.major_fao) > 0 # removes fishing inland
      AND nnet_score2 > .5
      AND seg_id IN (
        SELECT
          seg_id
        FROM
          good_segments)
      AND ssvid IN (
        SELECT
          ssvid
        FROM
          good_vessels
        WHERE
          year = 2018)
    GROUP BY
      year,
      ssvid,
      lat_bin_center,
      lon_bin_center,
      fao_region
    ORDER BY
      year,
      ssvid,
      fishing_hours_lat_lon DESC)

SELECT
  *,
  fishing_hours_lat_lon*engine_power_kw as fishing_KWh_lat_lon
FROM
  fishing_by_lat_lon
LEFT JOIN
  vessel_info
USING
  (ssvid,
  year)
'

binned_effort_table <- bq_table(project = bq_project, 
                                table = "effort_5deg_good_fishing_vessels_2018_raw",
                                dataset = "SubsidyExplorer")

if(bq_table_exists(binned_effort_table)){

  bq_table_delete(binned_effort_table)
  effort_binned_good_fishing_vessels_2018 <- 
    bq_project_query(bq_project, 
                    sql2, 
                    destination_table = binned_effort_table, 
                    use_legacy_sql = FALSE, 
                    allowLargeResults = TRUE) %>%
    bq_table_download()

}else {
  bq_table_create(binned_effort_table)
  effort_binned_good_fishing_vessels_2018 <- 
    bq_project_query(bq_project, 
                   query = sql2, 
                   destination_table = binned_effort_table,
                   use_legacy_sql = FALSE, 
                   allowLargeResults = TRUE) %>%
    bq_table_download()
}

# Save
write_csv(effort_binned_good_fishing_vessels_2018, paste0(results_dir, "effort_binned_good_fishing_vessels_2018_raw.csv"))
```

## By EEZ and/or FAO region

@@

```{r}
sql3 <- '
WITH
  # This subquery identifies good segments
  good_segments AS (
  SELECT
    seg_id
  FROM
    `world-fishing-827.gfw_research.pipe_v20190502_segs`
  WHERE
    good_seg
    AND positions > 10
    AND NOT overlapping_and_short),

@@

  bq_table_delete(effort_table)
  effort_good_fishing_vessels_2018 <- 
    bq_project_query(bq_project, 
                    sql3, 
                    destination_table = effort_table, 
                    use_legacy_sql = FALSE, 
                    allowLargeResults = TRUE) %>%
     bq_table_download()

}else {
  bq_table_create(effort_table)
  effort_good_fishing_vessels_2018 <- 
    bq_project_query("world-fishing-827", 
                   query = sql3, 
                   destination_table = effort_table,
                   use_legacy_sql = FALSE, 
                   allowLargeResults = TRUE) %>%
    bq_table_download()
}

# Save
write_csv(effort_good_fishing_vessels_2018, paste0(results_dir, "effort_good_fishing_vessels_2018_raw.csv"))
```
